Lesson 3
========================================================

Notes:


### Pseudo-Facebook User Data
Notes:
Using read.csv resulted in an file not found error

```{r Pseudo-Facebook User Data}
pf <- read.delim('pseudo_facebook.tsv')

```

### Histogram of Users' Birthdays

```{r Histogram of Users\' Birthdays}
# install.packages('ggplot2')
library(ggplot2)

names(pf)
qplot(x = dob_day, data = pf)
```
```{r}
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)
```
The same using ggplot

```{r}
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

### Faceting
Notes:

`facet_grid(vertical ~ horizontal)` can be used for multiple variables

```{r Faceting}
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month, ncol = 3)

```
```{r}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```



#### What code would you enter to create a histogram of friend counts?

```{r}
qplot(x = friend_count, data = pf)
```
```{r}
qplot(x = friend_count, data = pf, xlim = c(0, 1000))
```
Or Equivalently:
```{r}
qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0, 1000))
```
Or with `ggplot()`
```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000))
```

## Adjusting the Bin Width
```{r}
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

### Exploring with Bin Width
Notes:
```{r}
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
Remove NAs
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
With `ggplot()`
```{r}
ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)

```

# by() data a variable, a categorical variable and a function
```{r}
by(pf$friend_count, pf$gender, summary)
```

### Tenure
Notes:

```{r}
qplot(x = tenure, data = pf,
      color = I('black'), fill = I('#099DD9'))
```
```{r}
qplot(x = tenure, data = pf, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))
```
`ggplot()`
```{r}
ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

```{r}
ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
```
```{r}
qplot(x = tenure/365, data = pf,
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0,7))
```


### Labeling Plots

```{r}
qplot(x = tenure/365, data = pf,
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = pf, binwidth = 1,
      xlab = 'Age', ylab = 'Number of users',
      color = I('blue'), fill = I('green')) +
  scale_x_continuous(breaks = seq(0, 110, 5), lim=c(13,110))

```
In ggplot
```{r}
ggplot(aes(x = age), data = pf) +
  geom_histogram(binwidth = 1, fill = '#5760AB') +
  scale_x_continuous(breaks = seq(0, 113, 5))
```


### Transforming Data

Data sets with long tails should be transformed by eg. log to better
understand the data/turn it to a more normal distribution

```{r}
qplot(x = friend_count, data = pf)
summary(pf$friend_count)
```
```{r}
summary(log10(pf$friend_count))
```
We can't use log of 0 so:
```{r}
summary(log10(pf$friend_count + 1))
```

Another possible transformation is sqrt
```{r}
summary(sqrt(pf$friend_count + 1))
```

```{r}
library('gridExtra')
p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = log10(friend_count + 1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)

grid.arrange(p1, p2, p3, ncol = 1)
```
With ggplot()
```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```



### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
logScale <- qplot(x = log10(friend_count + 1), data = pf)

countScale <- ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_log10()

grid.arrange(logScale, countScale, ncol = 2)
```
```{r}
ggplot(aes(x = friend_count), data = pf) + 
  geom_histogram() +
  scale_x_log10()
```




### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), 
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```
```{r}
qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)),
      xlab = 'Friend Count', 
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```
In ggplot()
```{r}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```
### Likes on the Web

```{r}
qplot(x = www_likes, data=subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender) +
  scale_x_continuous()
```


This is long-tail data so let's use log()
```{r}
qplot(x = www_likes, data=subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender) +
  scale_x_continuous() +
  scale_x_log10()
```
Now we see at what point women overtake men in number of likes

```{r}
ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10()
```
```{r}
by(pf$www_likes, pf$gender, sum)
```



```{r}
qplot(x = www_likes, data=subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender) +
  scale_x_continuous() +
  scale_x_log10()
```


### Box Plots

```{r}
qplot(x = gender,, y = friend_count, 
      data=subset(pf, !is.na(gender)),
      geom = 'boxplot')
```
To ignore the outliers and focus on the box we can change the scale
```{r}
qplot(x = gender,, y = friend_count, 
      data=subset(pf, !is.na(gender)),
      geom = 'boxplot',
      ylim = c(0, 1000))
```
or
```{r}
qplot(x = gender,, y = friend_count, 
      data=subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0, 1000))

```
Since we have removed data from the set the boundaries may be incorrect, so...
```{r}
qplot(x = gender,, y = friend_count, 
      data=subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 1000))

```
Let's zoom in more
```{r}
qplot(x = gender,, y = friend_count, 
      data=subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 250))

```
Lets look at the numbers (remember by(variable to report, variable to split, function))
```{r}
by(pf$friend_count, pf$gender, summary)
```



#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)
```
The median of 4.0 shows that there are many 0 values
```{r}
summary(pf$mobile_likes > 0)
```
```{r}
mobile_checkin <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

```
What percent of check in using mobile?
```{r}
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```



Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!